---
title: "Lab1 244"
author: "Gage Clawson"
date: "1/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Objectives:

- basic review (data wrangling)
- intro to spatial data in R with sf
- intro to shiny apps

1. attach packages

```{r}
library(sf)
library(tidyverse)
library(leaflet)
library(tmap)
```

2. Review: reading, exploring, wrangling and visualizing data

a. read in the data
```{r}
wine <- read_csv("wine_data.csv")
```

b. wrangling

```{r}

wine_new <- wine %>%
  select(country, province, winery, region_1, points, price) %>%
  rename(state = province) %>%
  filter(state %in% c("California", "Oregon", "Washington")) %>% 
  mutate(ppd = points/price) %>% 
  arrange(-ppd) 

```

c. Use group_by and summarise to find summary stats by group

```{r}
wine_summary <- wine_new %>% 
  group_by(state) %>% 
  summarise(
    mean_ppd = mean(ppd, na.rm = TRUE),
    max_ppd = max(ppd, na.rm = TRUE), 
    min_ppd = min(ppd, na.rm = TRUE),
    count = length(ppd)
  )
```

d. A few gg graphics

```{r}

(wine_hist <- ggplot(wine_new, aes(x = ppd)) +
  geom_histogram(aes(fill = state)) +
   facet_wrap(~state, scales = "free")
 
 
 )
```

new graph: scatterplot graph of price vs points, with color determined by state, adding a linear model trend line on top of it

```{r}
wine_points <- ggplot(wine_new, aes(x = price, y = points)) +
  geom_point(aes(color = state), alpha = 0.5) +
  scale_color_manual(values = c("royalblue", "mediumorchid1", "orange")) +
  geom_smooth(method = "lm", aes(color = state), se = FALSE, size = 0.4) +
  facet_wrap(~state) +
  theme_light()

wine_points
```

2. California counties (spatial data) using sf

```{r}

ca_counties <- st_read(dsn = ".", layer = "CA_Counties_TIGER2016")

ca_land <- ca_counties %>% 
  select(NAME, ALAND)

ca_pop_inc <- read_csv("ca_pop_inc.csv") %>% 
  rename(NAME = COUNTY)

ca_df <- full_join(ca_land, ca_pop_inc) %>% 
  select(NAME, MedFamilyIncome)

# plot(ca_counties)



```

d. Use geom_sf to make a map in ggplot

```{r}

ca_income <- ggplot(ca_df) + 
  geom_sf(aes(fill = MedFamilyIncome))

ca_income

ca_df_transform <- st_transform(ca_df, crs = 4326)

leaflet(ca_df_transform) %>%
  addTiles() %>% #add bg
  addPolygons(weight = 1.0, 
              opacity = 1.0, 
              color = "white", 
              fillOpacity = 0.5,
              fillColor = ~colorQuantile("YlOrRd", MedFamilyIncome)(MedFamilyIncome))

```

f. ...or using tmap!

```{r}
tmap_mode("view")

if (Sys.getenv("USER") != "CRAN")

tm_shape(ca_df_transform) + tm_fill("MedFamilyIncome", alpha = 0.5)


```


###3. A Shiny app overview

a. Create a new Shiny app (single file containing ui and server). 
b. Notice that an example (Old Faithful Geyser histogram) appears. Select 'Run App' to run it - then play with the widget.

c. Follow along to create a new app (notice that a subfolder is created in WD for the app) for wine prices/points.










